Methods

### 2023-08-13 Analysis/clust_gmm.qmd
google sync

>>>>

```{r}
library(mixR)

mr1 <- mixfit(test$vx,ncomp=3)
plot(mr1)
mr1

mr2 <- mixfit(test$vx,ncomp=6,family="gamma")
plot(mr2)
mr2
```

library(flexmix)
#https://rdrr.io/cran/flexmix/man/flexmix.html

fm1=flexmix(vx~1|id,data=test,k=6)
summary(fm1)
plot(fm1)
parameters(fm1)

test %>% group_by(vb) %>% summarise(mean(vx))
```

```{r}
#http://www.sthda.com/english/articles/29-cluster-validation-essentials/96-determining-the-optimal-number-of-clusters-3-must-known-methods/

e1 <- readRDS(here("data/e1_08-04-23.rds"))
test <- e1 |> filter(expMode2 == "Test") |> ungroup()
library(factoextra)
library(NbClust)

fviz_nbclust(test |> select(vx), kmeans,method = "gap_stat")

NbClust(test |> select(vx),distance="euclidean",min.nc=2,max.nc=15,method="average")

```{r}
#| eval: false

library(brms)

testS <- test |> filter(id %in% 1:20)

options(mc.cores = 2, brms.backend = "cmdstanr")

mix <- mixture(gaussian, gaussian, gaussian,gaussian,gaussian,gaussian)
prior <- c(
  prior(normal(200, 100), Intercept, dpar = mu1),
  prior(normal(450, 100), Intercept, dpar = mu2),
  prior(normal(700, 100), Intercept, dpar = mu3),
  prior(normal(900, 100), Intercept, dpar = mu4),
  prior(normal(1100, 100), Intercept, dpar = mu5),
  prior(normal(1300, 100), Intercept, dpar = mu6)
)
fit1 <- brm(bf(vxC ~ 1), testS, family = mix,
            prior = prior, chains = 2)

summary(fit1)
pp_check(fit1,type="stat_grouped",ndraws=500,group="bandInt",stat="mean")

# saveRDS(object = fit1,
#         file = here("data/model_cache/mix6_vxC_test6.rda"), 
#         compress = "xz")

7x speedup for scaled vx
```{r}

library(brms)

test$scaleVx <- scale(test$vx)
test %>% group_by(vb) %>% summarise(mean(scaleVx))
# vb        `mean(scaleVx)`
#   <fct>               <dbl>
# 1 100-300           -0.544 
# 2 350-550           -0.273 
# 3 600-800           -0.0194
# 4 800-1000           0.449 
# 5 1000-1200          0.774 
# 6 1200-1400          1.00

testS <- test |> filter(id %in% 1:20)

options(mc.cores = 2, brms.backend = "cmdstanr")

mix <- mixture(gaussian, gaussian, gaussian,gaussian,gaussian,gaussian)
prior <- c(
  prior(normal(-.5, 10), Intercept, dpar = mu1),
  prior(normal(-.25, 10), Intercept, dpar = mu2),
  prior(normal(-.02, 10), Intercept, dpar = mu3),
  prior(normal(.4, 10), Intercept, dpar = mu4),
  prior(normal(.7, 10), Intercept, dpar = mu5),
  prior(normal(1, 10), Intercept, dpar = mu6)
)
fitScale <- brm(bf(scaleVx ~ 1), testS, family = mix,
            prior = prior, chains = 2)

summary(fitScale)
pp_check(fit1,type="stat_grouped",ndraws=500,group="bandInt",stat="mean")

# saveRDS(object = fit1,
#         file = here("data/model_cache/mix6_vxC_test6.rda"), 
#         compress = "xz")

options(mc.cores = 4, brms.backend = "cmdstanr")

mix <- mixture(gaussian, gaussian, gaussian,gaussian,gaussian,gaussian)
prior <- c(
  prior(normal(-.5, 10), Intercept, dpar = mu1),
  prior(normal(-.25, 10), Intercept, dpar = mu2),
  prior(normal(-.02, 10), Intercept, dpar = mu3),
  prior(normal(.4, 10), Intercept, dpar = mu4),
  prior(normal(.7, 10), Intercept, dpar = mu5),
  prior(normal(1, 10), Intercept, dpar = mu6)
)

fitScaleInt <- brm(bf(scaleVx ~ 1 + (1|id)), test, family = mix,
            prior = prior, chains = 2,iter=1000)

summary(fitScaleInt)
pp_check(fitScaleInt,type="stat_grouped",ndraws=500,stat="mean")

# saveRDS(object = fit1,
#         file = here("data/model_cache/mix6_vxC_test6.rda"), 
#         compress = "xz")
